\documentclass[acmsmall]{acmart}
\usepackage{tcolorbox}
\usepackage{fbox}
\usepackage{enumitem}
\usepackage{graphicx}
% \usepackage[table,xcdraw]{xcolor}
\usepackage{xcolor}

\usepackage{algorithm}
\usepackage{algpseudocode}

% for red strike
\usepackage{soul,xcolor}
\setstcolor{red}

\algrenewcommand\algorithmicensure{\textbf{Return:}}
\definecolor{ballblue}{rgb}{0.13, 0.67, 0.8}

\newcommand{\add}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}

\newcommand{\myauthornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\lp[1]{\myauthornote{LP}{ballblue}{#1}}
\newcommand\hh[1]{\myauthornote{HH}{purple}{#1}}
\newcommand\hy[1]{\myauthornote{HY}{red}{#1}}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}


% \acmBooktitle{Proc. ACM Softw. Eng. 1, FSE (FSE '24), (July, 2024)}

%%% The following is specific to FSE '24 and the paper
%%% 'BARO: Robust Root Cause Analysis for Microservices via Multivariate Bayesian Online Change Point Detection'
%%% by Luan Pham, Huong Ha, and Hongyu Zhang.
%%%
\setcopyright{rightsretained}
\acmDOI{10.1145/3660805}
\acmYear{2024}
\copyrightyear{2024}
\acmSubmissionID{fse24main-p752-p}
\acmJournal{PACMSE}
\acmVolume{1}
\acmNumber{FSE}
\acmArticle{98}
\acmMonth{7}
\received{2023-09-28}
\received[accepted]{2024-04-16}




\settopmatter{printacmref=true}
\pagestyle{headings} 

\begin{document}

\title{BARO: Robust Root Cause Analysis for Microservices via Multivariate Bayesian Online Change Point Detection}

\author{Luan Pham}
\orcid{0000-0001-7243-3225}
\affiliation{%
  \institution{RMIT University}
  % \city{Melbourne}
  \country{Australia}
}
\email{luan.pham@rmit.edu.au}

\author{Huong Ha}
\orcid{0000-0003-2463-7770}
\affiliation{%
  \institution{RMIT University}
  % \city{Melbourne}
  \country{Australia}
}
\email{huong.ha@rmit.edu.au}

\author{Hongyu Zhang}
\orcid{0000-0002-3063-9425}
\affiliation{%
  \institution{Chongqing University}
  % \city{Chongqing}
  \country{China}
}
\email{hyzhang@cqu.edu.cn}





\begin{abstract}

Detecting failures and identifying their root causes promptly and accurately is crucial for ensuring the availability of microservice systems. A typical failure troubleshooting pipeline for microservices consists of two phases: anomaly detection and root cause analysis. While various existing works on root cause analysis require accurate anomaly detection, there is no guarantee of accurate estimation with anomaly detection techniques. Inaccurate anomaly detection results can significantly affect the root cause localization results. To address this challenge, we propose \textit{BARO}, an end-to-end approach that integrates anomaly detection and root cause analysis for effectively troubleshooting failures in microservice systems. BARO leverages the Multivariate Bayesian Online Change Point Detection technique to model the dependency within multivariate time-series metrics data, enabling it to detect anomalies more accurately. BARO also incorporates a novel nonparametric statistical hypothesis testing technique for robustly identifying root causes, which is less sensitive to the accuracy of anomaly detection compared to existing works. Our comprehensive experiments conducted on three popular benchmark microservice systems demonstrate that BARO consistently outperforms state-of-the-art approaches in both anomaly detection and root cause analysis.



\end{abstract}
% \keywords{Root Cause Analysis, Microservices, AIOps}


%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10010940.10011003.10011004</concept_id>
       <concept_desc>Software and its engineering~Software reliability</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10010940.10011003.10011002</concept_id>
       <concept_desc>Software and its engineering~Software performance</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software reliability}
\ccsdesc[500]{Software and its engineering~Software performance}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Anomaly Detection, Root Cause Analysis, Microservice Systems}


\maketitle


\section{Introduction}
\label{sec:introduction}
\input{1.introduction.tex}


\section{Problem Statement and Background}
\label{sec:background}
\input{2.background}

\section{BARO: Proposed Root Cause Analysis Approach}
\label{sec:method}
\input{3.method.tex}

\section{Evaluation}

This section answers the following research questions:
\begin{itemize}
    \item \textbf{\textit{RQ1: How effective is BARO in anomaly detection?}} To answer this RQ, we conduct an experiment to compare BARO with state-of-the-art anomaly detection approaches and evaluate their performance in detecting anomalies.
    \item \textbf{\textit{RQ2: How effective is BARO in root cause analysis?}} To answer this RQ, we compare BARO with state-of-the-art RCA methods and evaluate their performance in ranking both root cause services (coarse-grained) and root cause metrics (fine-grained) of the failures.
    \item \textit{\textbf{RQ3: How effective are the components of BARO?}} To answer this RQ, we evaluate the effectiveness of each main component of BARO: Multivariate BOCPD in detecting anomalies and RobustScorer in locating the failure's root cause.
    \item \textbf{\textit{RQ4: How sensitive different RCA methods are w.r.t different parameters?}} To answer this RQ, we perform a sensitivity analysis to evaluate the peformance of all these methods with different values of the anomaly detection time and other methods hyperparameters.
\end{itemize}
    
\label{sec:results}
\input{4.results.tex}

\label{sec:discussion}
\input{5.discussion.tex}

\section{Conclusion}
\label{sec:conclusion}

This paper proposes BARO, a novel end-to-end approach for anomaly detection and root cause analysis for microservice systems based on multivariate time series metrics data. BARO uses Multivariate Bayesian Online Change Point Detection for anomaly detection and introduces the RobustScorer, a nonparametric statistical hypothesis testing technique, for identifying root causes. Our experimental results on three benchmark systems demonstrate that BARO consistently outperforms the state-of-the-art methods in both anomaly detection and root cause analysis. Comprehensive sensitivity analysis highlights the robustness and broad applicability of our approach. Our research contributes a valuable tool for metric-based root cause analysis of microservice systems. 
In future work, we plan to enhance our framework by incorporating multimodal data (e.g., logs, traces). 
% In future work, we plan to enhance anomaly detection and RCA by incorporating multimodal data (e.g., logs, traces) and integrating new advances in ML/DL (e.g., large language models). 
% We also plan to enhance the RCA's trustworthiness by providing explainable root causes.

\section{Data Availability}
We have open-sourced BARO, which can be accessed on GitHub at \cite{barogithub}. Additionally, an immutable artifact for BARO is available on Zenodo \cite{barogithubzenodo}, together with three experimental datasets \cite{barodatasets}.


% The replication package for this paper, including the source code of BARO, three experimental datasets, and the reproducibility instructions, are publicly available \cite{baroartifacts}. 


\begin{acks}
This research was supported by the Australian Research Council Discovery Project (DP220103044), AWS Cloud Credit for Research, and Google Travel Grant. We also thank anonymous reviewers for their insightful and constructive comments, which significantly improve this paper. 
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{reference}

\end{document}
