\documentclass[acmsmall,nonacm]{acmart} % remove "nonacm" to print all ACM stuff

\usepackage{array}
\usepackage{cleveref}
\usepackage{colortbl}
\usepackage{csquotes}
\usepackage{mdframed}
\usepackage{xspace}

%% These commands are for a JOURNAL article.
\acmJournal{CSUR}
%\acmVolume{1}
%\acmNumber{1}
%\acmArticle{1}
%\acmMonth{1}

%% Config
\hypersetup{
  pdftitle = {Anomaly Detection and Root Cause Analysis - Survey},
  pdfauthor = {Soldani J. and Brogi A.}
}

%% MACROS - Utility
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\figscale}{.3}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\viz}{viz.,\xspace}

\newcommand{\mycomment}[3]{\smallskip\begin{mdframed}[backgroundcolor=#3]\textbf{#1:} #2\end{mdframed}}
\newcommand{\note}[1]{\mycomment{NOTE}{#1}{yellow!20}}
\newcommand{\tbd}[1]{\mycomment{TBD}{#1}{gray!20}}

\newcommand{\upd}[1]{#1} %{\color{magenta} #1}}

%% MACROS - Table
\newcommand{\thickline}{\specialrule{1pt}{0.2pt}{0.2pt}}
\newcommand{\thinline}{\specialrule{.4pt}{0.1pt}{0.1pt}}

\begin{document}

\title{Anomaly Detection and Failure Root Cause Analysis in (Micro)Service-Based Cloud Applications: A Survey}

\renewcommand{\shorttitle}{Anomaly Detection and Failure Root Cause Analysis in (Micro)Service-Based Cloud Applications}

\author{Jacopo Soldani}
\email{jacopo.soldani@unipi.it}
\author{Antonio Brogi}
\email{antonio.brogi@unipi.it}
\affiliation{%
  \institution{University of Pisa}
  \streetaddress{Dipartimento di Informatica, Largo B. Pontecorvo 3}
  \city{Pisa}
  \state{PI}
  \country{Italy}
  \postcode{56127}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{J. Soldani and A. Brogi}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
The momentum gained by microservices and cloud-native software architecture pushed nowadays enterprise IT towards multi-service applications.
The proliferation of services and service interactions within applications, 
% often being composed by hundreds of interacting services, makes it harder to detect failures and identify their possible root causes, which is however crucial to promptly recover and fix applications.
% Various techniques have hence been proposed to promptly detect failures based on their symptoms, \viz observing anomalous behaviour in one or more application services, as well as to analyse the logs or monitored performance of such services to support application operators in determining the possible root causes for observed anomalies.
% This survey provides a structured overview of currently available techniques for anomaly detection and root cause analysis in modern multi-service applications, with the aim of supporting application operators in choosing the techniques most suited to their applications' needs.
% The survey also qualitatively analyses the surveyed techniques under multiple dimensions, \eg the need for a tradeoff between their setup costs and accuracy, and it elaborates on the open challenges and research directions stemming out from the analysis.
often consisting of hundreds of interacting services, makes it harder to detect failures and to identify their
possible root causes, which is on the other hand crucial to promptly recover and fix applications. Various techniques
have been proposed to promptly detect failures based on their symptoms, viz., observing anomalous
behaviour in one or more application services, as well as to analyse logs or monitored performance of such
services to determine the possible root causes for observed anomalies. 
%
The objective of this survey is to provide a structured overview and a qualitative analysis of currently available techniques for anomaly detection and root cause analysis in modern multi-service applications.
Some open challenges and research directions stemming out from the analysis are also discussed.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
% \begin{CCSXML}
% <ccs2012>
%    <concept>
%        <concept_id>10010405.10010406.10010421</concept_id>
%        <concept_desc>Applied computing~Service-oriented architectures</concept_desc>
%        <concept_significance>300</concept_significance>
%        </concept>
%    <concept>
%        <concept_id>10002944.10011122.10002945</concept_id>
%        <concept_desc>General and reference~Surveys and overviews</concept_desc>
%        <concept_significance>100</concept_significance>
%        </concept>
%  </ccs2012>
% \end{CCSXML}

% \ccsdesc[300]{Applied computing~Service-oriented architectures}
% \ccsdesc[100]{General and reference~Surveys and overviews}

%%
%% Keywords. The author(s) should pick words that accurately describe the work being presented. Separate the keywords with commas.
% \keywords{microservices, multi-service applications, failure detection, anomaly detection, root cause analysis}


%%
%% This command processes the author and affiliation and title information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
\label{sec:intro}
\input{src/1-intro.tex}

\smallskip \noindent
The rest of this article is organised as follows.
\Cref{sec:terminology} introduces the terminology used in our survey.
\Cref{sec:detection,sec:rca} provide a structured overview of the existing techniques for detecting anomalies in multi-service applications and for identifying their root causes, respectively.
\Cref{sec:related} positions our survey with respect to related work.
\Cref{sec:conclusions} concludes the article by discussing possible research directions on detecting and explaining anomalies in multi-service applications.

\section{Terminology}
\label{sec:terminology}
\input{src/2-terminology.tex}

\section{Anomaly Detection}
\label{sec:detection}
\input{src/3-detection}

\section{Root Cause Analysis}
\label{sec:rca}
\input{src/4-root-cause}

\section{Related Work}
\label{sec:related}
\input{src/5-related.tex}

\section{Conclusions}
\label{sec:conclusions}
\input{src/6-conclusions.tex}


%%
%% The next two lines define the bibliography style to be used, and the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{src/biblio}

%%
%% If your work has an appendix, this is the place to put it.
%\appendix

\end{document}
\endinput
%%
%% End of file `sample-acmsmall.tex'.
