%\documentclass[conference]{IEEEtran}
\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{T1}{fontenc}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{makecell}
\usepackage{subfigure}
\usepackage{color}
\usepackage{soul}
\newcommand{\red}[1]{\textcolor{red}{#1}}

\newcommand\app{MicroHECL}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{MicroHECL: High-Efficient Root Cause Localization in Large-Scale Microservice Systems}

\author{
\IEEEauthorblockN{\hspace{0.14cm}Dewei Liu\hspace{1cm}}
\IEEEauthorblockA{
\textit{\hspace{0.14cm}Fudan University\hspace{1cm}}\\
\hspace{0.14cm}Shanghai, China\hspace{1cm}}
\and
\IEEEauthorblockN{Chuan He\hspace{1cm}}
\IEEEauthorblockA{
\textit{Fudan University\hspace{1cm}}\\
Shanghai, China\hspace{1cm}}
\and
\IEEEauthorblockN{Xin Peng\hspace{1cm}}
\IEEEauthorblockA{
\textit{Fudan University\hspace{1cm}}\\
Shanghai, China\hspace{1cm}}
\and
\IEEEauthorblockN{Fan Lin\hspace{1cm}}
\IEEEauthorblockA{
\textit{Alibaba Group\hspace{1cm}}\\
Hangzhou, China\hspace{1cm}}
\and
\IEEEauthorblockN{Chenxi Zhang\hspace{1cm}}
\IEEEauthorblockA{
\textit{Fudan University\hspace{1cm}}\\
Shanghai, China\hspace{1cm}}
\and
\IEEEauthorblockN{\hspace{1cm}Shengfang Gong\hspace{1cm}}
\IEEEauthorblockA{
\textit{\hspace{1cm}Alibaba Group\hspace{1cm}}\\
\hspace{1cm}HangZhou, China\hspace{1cm}}
\and
\IEEEauthorblockN{Ziang Li\hspace{1cm}}
\IEEEauthorblockA{
\textit{Alibaba Group\hspace{1cm}}\\
HangZhou, China\hspace{1cm}}
\and
\IEEEauthorblockN{Jiayu Ou\hspace{1cm}}
\IEEEauthorblockA{
\textit{Alibaba Group\hspace{1cm}}\\
HangZhou, China\hspace{1cm}}
\and
\IEEEauthorblockN{Zheshun Wu\hspace{1cm}}
\IEEEauthorblockA{
\textit{Alibaba Group\hspace{1cm}}\\
HangZhou, China\hspace{1cm}}
}

\maketitle

\begin{abstract}
Availability issues of industrial microservice systems (e.g., drop of successfully placed orders and processed transactions) directly affect the running of the business.
These issues are usually caused by various types of service anomalies which propagate along service dependencies.
Accurate and high-efficient root cause localization is thus a critical challenge for large-scale industrial microservice systems.
Existing approaches use service dependency graph based analysis techniques to automatically locate root causes.
However, these approaches are limited due to their inaccurate detection of service anomalies and inefficient traversing of service dependency graph.
In this paper, we propose a high-efficient root cause localization approach for availability issues of microservice systems, called \app.
Based on a dynamically constructed service call graph, \app~analyzes possible anomaly propagation chains, and ranks candidate root causes based on correlation analysis.
We combine machine learning and statistical methods and design customized models for the detection of different types of service anomalies (i.e., performance, reliability, traffic).
To improve the efficiency, we adopt a pruning strategy to eliminate irrelevant service calls in anomaly propagation chain analysis.
Experimental studies show that \app~significantly outperforms two state-of-the-art baseline approaches in terms of both accuracy and efficiency.
\app~has been used in Alibaba and achieves a top-3 hit ratio of 68\% with root cause localization time reduced from 30 minutes to 5 minutes.


\end{abstract}

\begin{IEEEkeywords}
microservice, availability, root cause localization, anomaly detection, service call graph
\end{IEEEkeywords}

\section{Introduction}
\input{introduction}

\section{Background}
\input{background}

\section{\app~Overview}\label{sec:overview}
\input{overview}

\section{Anomaly Propagation Chain Analysis}\label{sec:propagation}
\input{propagation}

\section{Experimental Study}
\input{experiment}

\section{Practical Application}
\input{application}

\section{Related Work}
\input{related}

\section{Conclusion}
\input{conclusion}


\bibliographystyle{IEEEtran}
\bibliography{reference}
\end{document}
